Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-09-22 Ben_Manuscripts/Paper1/Outline.tex
small change
addressed an issues compiling
organizing

  We tested a number of equilibration schemes in an effort to match experiment
  while overcoming kinetic limitations of such a large system that can easily
  get stuck in multiple metastable states.
    Equilibration Scheme A:
    \item Apply position restraints to aromatic rings in order to maintain
    orientational order. 
    \item A short, restrained equilibration, followed by NPT simulations 
    between 400 and 500 ns, allows the initial configuration to relax into
    an equilibrium configuration.
    \item Our logic for choosing a starting configuration and the details 
    of the equilibration schemes are presented below.
  \end{itemize}

@@

  We were inspired to ask question (4) when we observed that the initial distance
  between layers can influence the approach towards equilibrium. 
  \begin{itemize}
  	\item Our choice of distance between layers results in long-term stability of 
	different configurations indicating that we have found more than one metastable
	free energy basin.
  	\item Equilibrated systems built according to the 3.7 \angstrom~ 
	layer spacing implied by R-$\pi$ are characterized by a defined, cylindrical and
	open pore structure.
        \item We will refer to this large set of configurations, with an open pore, as 
        Basin OP (open pore) (Figure~\ref{fig:abasin}).
	\item Simulations of systems built with layers stacked greater than 4 
	\angstrom~ apart results in a pore structure characterized by high radial 
	disorder, while still maintaining partitioning between hydrophobic and 
	hydrophilic regions.
	\item This will be called Basin CP (closed pore) (Figure~\ref{fig:bbasin}).
	\item This LLC membrane may exist in at least two metastable states.
	\item The distinct difference in pore structure exhibited by each phase will 
	likely lead to different transport mechanisms.
	\item Because the pore structure will vary between each Basin, understanding 
	which phase exists experimentally is necessary in order to ensure we are 
	studying the system which actually dominates.
  \end{itemize}

@@

  %MRS: one thought: maybe organize it a different way -- it's not the disordered (and go through the anlysis there), and then separately show that it's more likely the sandwich.  Something to think about.
  %MRS: anyway, this is the key part of the paper, so clarity and organization important.
  %MRS: make sure the figures are comparable.  Could someone look at the figures, and see immediately what you are explaining -- for example, experiment and fully equilibrated disordered, sandwich and parallel displaced all in a row?
  Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  patterns produced from equilibrated MD trajectories shows the most consistency
  with the offset configuration in the OP basin.
  \begin{itemize}
  	\item Systems were built in both the parallel displaced and sandwiched 
	configurations with an intial layer spacing of 3.7 \angstrom~.
	\item A third system was created by stacking layers in the sandwiched 
	configuration 5 \angstrom~apart in order to guide it towards the CP Basin.
	% BJC: I've done an additional offset disordered system which goes to Basin CP but I think that will confuse the point
	\item For ease of reference, the third system will be referred to as Basin D (disordered). 
%BJC: omit and call it basin B
%MRS or just be consistent from the beginning.  Or call Basins O and D          
	\item The three systems were equilibrated according to our procedure with
	NPT simulations of greater than 400 ns.
	\item Simulated diffration patterns were generated using portions of the
	trajectory after equilibration.  
	\item We assume the simulation has equilibrated when the distance between 
	pores and the membrane thickness stopped changing.
	\item Simulated diffraction patterns for all three structures are shown in 
	Figure ~\ref{fig:XRDsim}. 
	%BJC: maybe break into a separate paragraph starting here
	\item The equilibrated parallel displaced structure exhibits R-alkanes and
	R-pores in the expected locations. 
	\item Due to low resolution, making out the individual spots of R-pores is
	challenging and can be validated upon full spherical integration of the 3D 
	structure factor. However, the same information can be extracted by measuring 
	the pore spacing as described earlier.
	\item R-$\pi$ is also present, intersecting R-alkanes at a lower q value than
	in experiment. The rings prefer to stack $\approx 4.1 \angstrom$ apart as 
	opposed to 3.7 \angstrom~.
	\item R-helix and R-spots are also both missing.
	\item The line which intersected both of these features in the restrained 
	simulations faded upon equilibration which suggests that monomers in the 
	parallel-displaced configuration may not give rise to the experimental WAXS 
	pattern seen.
	\item Simulated XRD of the sandwiched configuration contains all features 
	except for R-helix.
	\item Most notably, R-spots appears in the expected location, which suggests 
	that there is something intrinsic to the sandwiched packing that gives rise 
	to such features. 
	\item Finally, the disordered structure exhibits R-alkanes and R-pores but 
	R-helix, R-$\pi$ and R-spots are not present.
  \end{itemize}

@@

  The spots that appear in the simulated XRD pattern of OP basin
  conformations are a result of the way alkane tails pack together.
  \begin{itemize}
  	\item Previously, the spots in the diffraction pattern had been explained 
	as the product of tilted alkane chains.
	\item We measured the tilt angle of the alkane chains and showed that our 
	system equilibrates to an average tilt angle close to zero degrees (See Supplemental Information). 
	\item To understand the origin of the spots, we determined which atoms gave rise to the feature.
	\item By removing atoms from the trajectory and simulating a diffraction 
	pattern, we were able to isolate the cause of the spots to the tails (Figure~\ref{fig:tails_rzplot}).
	\item Since the tails stay nearly flat, we plotted the centroids of the 
	tails and measured the angle between each centroid and its nearest neighbors
	with respect to the plane of the membrane (Figure~\ref{fig:centroids}).
	\item The distribution of these angles is consistent with the location of 
	the spots (Figure~\ref{fig:angle_distribution}).
	\item We reason that monomer tails stacked closely in the sandwiched conformation
	are forced to splay apart and pack in between in each other which creates a
	nearly hexagonal array of packed tails. 
  \end{itemize}

  \begin{figure}[!ht]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\hspace{-1cm}
		\vspace{1cm}
		\includegraphics[width=\textwidth,scale=2]{tails_topview.png}
		\caption{}\label{fig:tails_topview}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{tails_rzplot.png}
		\caption{}\label{fig:tails_rzplot}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{centroids_box.png}
		\caption{}\label{fig:centroids}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{angles_traj_layered.png}
		\caption{}\label{fig:angle_distribution}
	\end{subfigure}
	\caption{(a) The trajectory can be stripped of all atoms except carbon
	atoms in monomer tails. (b) Simulated diffraction of the tail-only trajectory
	still gives rise to R-spots. (c) Finding the center of mass and visualizing 
	their coordinates reveals the hexagonal-like packing of the tails. (d) The 
	distribution created by measuring the angle between each centroid (e.g. red 
	in (c)) and its neareset neighbors (e.g. green in (c)) with respect to the xy
	plane has distinct spikes near 30\degree, which is consistent with the location
	of R-spots}\label{fig:tail_packing}
  \end{figure}
  %MRS: one could ask the question about basins for the parallel displaced and sandwich as well.  
  %MRS: we had talked a lot about interconversion (or conversion one way but not other?) but 
  %MRS: I think things have changed since then? 
  We can not immediately classify Basin A and Basin B as separate phases.
  \begin{itemize}
	\item To prove the existence of two phases we need evidence of a
	a first order phase transition.
	\item A first order phase transition can be denoted by a discontinuity
        of some order parameter in response to an external condition such as
	temperature.
	\item We chose three easily measurable order parameters: the distance
	between pores, the membrane thickness and the ratio of pore radius to
	the uncertainty in pore radius.  % BJC: these may change
        \item The pore radius is divided by its uncertainty as a way of quantifying
        the degree to which monomers obstruct the pore region.
  \end{itemize}

  Basin B is the dominant configuration at higher temperatures.
  \begin{itemize}
	\item We linearly ramped the temperature of a system in Basin A 
	from 280K to 340K over 100 ns.
        \item Visually, there is a distinct change in pore structure from one
        characteristic of Basin A (Figure~\ref{fig:280K_pore}) to one characteristic of
        Basin B (Figure~\ref{fig:340K_pore}).
        \item The slope of all order parameters changes between 315K and 325K
	(\Cref{fig:p2p_tramp,fig:thickness_tramp,fig:order_tramp}) indicating the
        possibility of an abrupt change in system ordering.
        \item Our 100 ns temperature ramp was likely too fast and caused the system
        to suffer from hysteresis.
  \end{itemize}

@@

  We have used a detailed molecular model of the Col\subscript{h} phase formed by NA-GA3C11
  in order to study the nanoscopic structure.
  \begin{itemize}
  	\item While there have been efforts to model formation of various liquid crystalline 
	phases with molecular dynamics, to our knowledge there have been no studies which 
	attempt to examine their structure with the same level of detail presented here.
	\item We have confirmed that monomers stay partitioned in layers.
	\item We were able to deduce that layers are composed of 5 monomers.
	\item We have identified two metastable basins which each consist of a set of 
	similar monomer configurations characterized by the degree of order in the pore region.
	\item We verified that the basins are not separate phases.
	\item We have explored the affect of two different $\pi$-$\pi$ stacking modes on the equilibrated
	membrane strucure. 
	\item Simulated diffration patterns generated from MD trajectories suggest that
	the offset configuration produces a structure with the closest match to experiment. 
	\item Even though our model only answers these questions for a system created by 
	NA-GA3C11, it can be adapted to other study systems formed by other LCs with 
	little extra effort. 
  \end{itemize}

>>>>

  We tested a number of equilibration schemes in an effort to match experiment
  while overcoming kinetic limitations of such a large system that can easily
  get stuck in multiple metastable states.
    Equilibration Scheme A:
  \begin{itemize}
    \item Apply position restraints to aromatic rings in order to maintain
    orientational order. 
    \item A short, restrained equilibration, followed by NPT simulations 
    between 400 and 500 ns, allows the initial configuration to relax into
    an equilibrium configuration.
    \item Our logic for choosing a starting configuration and the details 
    of the equilibration schemes are presented below.
  \end{itemize}

@@

  In addition to the parallel displaced and sandwiched configurations we observed
  that the initial distance between layers can influence the approach towards 
  equilibrium.
  \begin{itemize}
  	\item Our choice of distance between layers results in long-term stability of 
	qualitatively different configurations suggesting that we have found more than
        one metastable free energy basin.
  	\item Equilibrated systems built according to the 3.7 \angstrom~layer spacing 
        implied by R-$\pi$ are characterized by a defined, cylindrical and open pore 
	structure.
        \item We will refer to this large set of configurations, with an open pore, as 
        Basin OP (open pore) (Figure~\ref{fig:abasin}).
	\item Simulations of systems built with layers stacked greater than 4 
	\angstrom~ apart results in a pore structure characterized by high radial 
	disorder, while still maintaining partitioning between hydrophobic and 
	hydrophilic regions.
	\item This will be called Basin CP (closed pore) (Figure~\ref{fig:bbasin}).
	\item This LLC membrane may exist in at least two metastable states.
	\item The distinct difference in pore structure exhibited by each phase will 
	likely lead to different transport mechanisms.
	\item Because the pore structure will vary between each Basin, understanding 
	which phase exists experimentally is necessary in order to ensure we are 
	studying the system which actually dominates.
  \end{itemize}

@@

  %MRS: one thought: maybe organize it a different way -- it's not the disordered (and go through the anlysis there), and then separately show that it's more likely the sandwich.  Something to think about.
  %MRS: anyway, this is the key part of the paper, so clarity and organization important.
  %MRS: make sure the figures are comparable.  Could someone look at the figures, and see immediately what you are explaining -- for example, experiment and fully equilibrated disordered, sandwich and parallel displaced all in a row?

  %Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  %patterns produced from equilibrated MD trajectories shows the most consistency
  %with the offset configuration in the OP basin.

  We created systems using three 
  distinct initial configurations and examined their structure by simulating 
  X-ray diffraction patterns produced from equilibrated MD trajectories.
  \begin{itemize}
  	\item Systems were built in both the parallel displaced and sandwiched 
	configurations with an intial layer spacing of 3.7 \angstrom~.
	\item A third system was created by stacking layers in the sandwiched 
	configuration 5 \angstrom~apart in order to guide it towards the CP Basin.
	% BJC: I've done an additional offset disordered system which goes to Basin CP but I think that will confuse the point
	\item For ease of reference, the third system will be referred to as Basin D (disordered). 
	\item The three systems were equilibrated according to our procedure with
	NPT simulations of greater than 400 ns.
	\item Simulated diffration patterns were generated using portions of the
	trajectory after equilibration.  
	\item We assume the simulation has equilibrated when the distance between 
	pores and the membrane thickness stopped changing.
	\item Simulated diffraction patterns for all three structures are shown in 
	Figure ~\ref{fig:XRDsim}. 
  \end{itemize}

  % Three paragraphs instead here? Talk about disordered, parallel displaced and 
  % layered separately? Try it. 
  % Talk about disordered last so it leads in to 
  % Also maybe don't talk about question 4 until after this
  Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  patterns shows the most consistency with the offset configuration in the OP basin.
  \begin{itemize}
        \item The disordered pore structure exhibits R-alkanes and R-pores but R-helix,
	R-$\pi$ and R-spots are not present. 
        \item Simulated XRD of the sandwiched configuration contains all features 
        except for R-helix.
        \item Most notably, R-spots appears in the expected location, which suggests 
        that there is something intrinsic to the sandwiched packing that gives rise 
        to such features.

	\item The equilibrated parallel displaced structure exhibits R-alkanes and
	R-pores in the expected locations. 
	\item Due to low resolution, making out the individual spots of R-pores is
	challenging and can be validated with slightly better resolution upon full
	spherical integration of the 3D structure factor. However, the same 
	information can be extracted by measuring the pore spacing as described earlier.
	\item R-$\pi$ is also present, intersecting R-alkanes at a lower q value than
	in experiment. The rings prefer to stack $\approx 4.1 \angstrom$ apart as 
	opposed to 3.7 \angstrom~.
	\item R-helix and R-spots are also both missing.
	\item The line which intersected both of these features in the restrained 
	simulations faded upon equilibration which suggests that monomers in the 
	parallel-displaced configuration may not give rise to the experimental WAXS 
	pattern seen.
	\item Simulated XRD of the sandwiched configuration contains all features 
	except for R-helix.
	\item Most notably, R-spots appears in the expected location, which suggests 
	that there is something intrinsic to the sandwiched packing that gives rise 
	to such features. 
  \end{itemize}

@@

  The spots that appear in the simulated XRD pattern of OP basin
  conformations are a result of the way alkane tails pack together.
  \begin{itemize}
  	\item Previously, the spots in the diffraction pattern had been explained 
	as the product of tilted alkane chains.
	\item We measured the tilt angle of the alkane chains and showed that our 
	system equilibrates to an average tilt angle close to zero degrees (See Supplemental Information). 
	\item To understand the origin of the spots, we determined which atoms gave rise to the feature.i
	\item Since R-spots is present as higher intensity spots within R-alkanes, it is likely
        that the spots arise as a consequence of the tails. 
	\item By removing atoms from the trajectory and simulating a diffraction 
	pattern, we were able to isolate the cause of the spots to the tails (Figure~\ref{fig:tails_rzplot}).
	\item Since the tails stay nearly flat, we plotted the centroids of the 
	tails and measured the angle between each centroid and its nearest neighbors
	with respect to the plane of the membrane (Figure~\ref{fig:centroids}).
	\item The distribution of these angles is consistent with the location of 
	the spots (Figure~\ref{fig:tail_packing}).
	\item The peaks of interest in Figures~\ref{fig:offset_tails}~and~\ref{fig:layered_tails}
	are located at $\pm$ 33 $\degree$ which is the same location where the highest intensity
	of spots are located on the simulated patterns (See Supplemental Information for quantitative proof)
	\item We integrated the raw experimental 2D WAXS data in the region bounding R-alkanes 
        and found the angle at which R-spots reaches its highest intensity to be $\pm$ 37 $\degree$ which 
        is a reconcilable difference with our simulated results.  
  \end{itemize}

  \begin{figure}[ht]
  \centering
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{offset_tail_packing.png}
        \caption{}~\label{fig:offset_tails}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{angles_traj_layered.png}
        \caption{}~\label{fig:layered_tails}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{integrated_WAXS_ring.png}
        \caption{}~\label{fig:rz_layered}
  \end{subfigure}
  \caption{The distribution of angles w.r.t. xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (b) configurations both show
  distinct peaks at $\pm$ 33 $\degree$. (c) Integrated 2D WAXS data between $q_r$=1 and $q_r$=1.42 shows
  distinct peaks at with maxima at $\pm$ 37 $\degree$}~\label{fig:tail_packing}
  \end{figure}

  %MRS: one could ask the question about basins for the parallel displaced and sandwich as well.  
  %MRS: we had talked a lot about interconversion (or conversion one way but not other?) but 
  %MRS: I think things have changed since then? 
  The disordered basin shares little in common with the ordered basin but 
  its long term stability suggests that it can exist under some conditions.  
  Basin B is the dominant configuration at higher temperatures.
  \begin{itemize}
        \item We linearly ramped the temperature of a system in Basin A
        from 280K to 340K over 100 ns.
        \item Visually, there is a distinct change in pore structure from one
        characteristic of Basin A (Figure~\ref{fig:280K_pore}) to one characteristic of
        Basin B (Figure~\ref{fig:340K_pore}).
        \item The slope of all order parameters changes between 315K and 325K
        (\Cref{fig:p2p_tramp,fig:thickness_tramp,fig:order_tramp}) indicating the
        possibility of an abrupt change in system ordering.
        \item Our 100 ns temperature ramp was likely too fast and caused the system
        to suffer from hysteresis.
  \end{itemize}

  We can not immediately classify Basin A and Basin B as separate phases.
  \begin{itemize}
	\item To prove the existence of two phases we need evidence of a
	a first order phase transition.
	\item A first order phase transition can be denoted by a discontinuity
        of some order parameter in response to an external condition such as
	temperature.
	\item We chose three easily measurable order parameters: the distance
	between pores, the membrane thickness and the ratio of pore radius to
	the uncertainty in pore radius.  % BJC: these may change
        \item The pore radius is divided by its uncertainty as a way of quantifying
        the degree to which monomers obstruct the pore region.
  \end{itemize}

@@

  We have used a detailed molecular model of the Col\textsubscript{h} phase formed by NA-GA3C11
  in order to study the nanoscopic structure.
  \begin{itemize}
  	\item While there have been efforts to model formation of various liquid crystalline 
	phases with molecular dynamics, to our knowledge there have been no studies which 
	attempt to examine their structure with the same level of detail presented here.
	\item We have confirmed that monomers stay partitioned in layers.
	\item We were able to deduce that layers are composed of 5 monomers.
	\item We have identified two metastable basins which each consist of a set of 
	similar monomer configurations characterized by the degree of order in the pore region.
	\item We verified that the basins are not separate phases.
	\item We have explored the affect of two different $\pi$-$\pi$ stacking modes on the equilibrated
	membrane strucure. 
	\item Simulated diffration patterns generated from MD trajectories suggest that
	the offset configuration produces a structure with the closest match to experiment. 
	\item Even though our model only answers these questions for a system created by 
	NA-GA3C11, it can be adapted to other study systems formed by other LCs with 
	little extra effort. 
  \end{itemize}
